Natural sources of vagueness and their implications

### 2017-02-15 book_chapter/paper.tex
Added introduction to signaling games.

\usepackage{todonotes}
\usepackage{a4wide}
\usepackage{amssymb}
\usepackage{syllogism}

@@

>>>>

\usepackage{todonotes}
\usepackage{a4wide}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{syllogism}

@@

Signaling games were first introduced as models of communication by David Lewis~\parencite*{lewis_convention_1969}.
He wanted to address an argument raised by W.~V.~Quine~\parencite*{quine_truth_1936} and others against the possibility of language having had started as a conventional system: if language is a convention, it had to be originally established by an agreement; in order to establish an agreement, a convention-governed system of communication would have to already have been in place; thus, although some languages could have been established by agreement if another convention was already in place, not all of them could.
To this, Lewis retorts:
\begin{quote}
I offer this rejoinder: an agreement sufficient to create a convention need not be a transaction involving language or any other conventional activity.
All it takes is an exchange of manifestations of a propensity to conform to a regularity.%
~\parencite*[87--88]{lewis_convention_1969}
\end{quote}
In order to support this claim, Lewis studies coordination problems formalized in terms of game theory.

Coordination problems are ``situations of interdependent decision by two or more agents in which coincidence of interest predominates and in which there are two or more proper coordination equilibria''~\parencite*[24]{lewis_convention_1969}.
In game theory terms, the agents interested in the coordination are the players, the game involves each player making an independent choice from his set of available actions, a payoff is what is attributed to each player based on the choices of both.
The setup of the game, \emph{i.e.}~the available actions and the relative interests of the players, can be represented in a payoff matrix, where rows show one player's available actions, columns the other player's available actions, and each cell gives the payoff for each of the players based on its row and column combination.
For example, consider the following matrix:
\begin{center}
\begin{tabular}{r|c|c|}
\multicolumn{1}{r}{}
 & \multicolumn{1}{c}{$b_1$}
 & \multicolumn{1}{c}{$b_2$} \\ \cline{2-3}
   $a_1$ & $1,1$ & $0,0$ \\ \cline{2-3}
   $a_2$ & $0,0$ & $1,1$ \\ \cline{2-3}
\end{tabular}
\end{center}
We can see this as representing the following game: one player has actions $A = \lbrace a_1, a_2 \rbrace$ available, the other $B = \lbrace b_1, b_2 \rbrace$; they prefer to coordinate $a_1$ with $b_1$ or $a_2$ with $b_2$, thus if this is achieved each gets a payoff of $1$, otherwise they each get $0$.
Formally, all we need to define in order to characterize such a situation are the sets of actions $A$ and $B$, and a utility function $U : A \times B \rightarrow \mathbb{R}^2$ that specifies the payoff for each player given the choices of actions.

The above example is a very simple case, but it serves to illustrate Lewis' notion of convention.
The two pairs of choices $(a_1,b_1)$ and $(a_2,b_2)$ are stable coordination equilibria, because in such a scenario no player has an incentive to unilaterally change his choice: if the first player is going to choose action $a_1$, the second player would get a payoff of $0$ for switching to $b_2$, instead of $1$ from sticking to $b_1$; the same reasoning applies, mutatis mutandis, to the other player and the other equilibrium.
However, neither player has a preference between those two pairs.
Furthermore, if the same coordination problem arises again, Lewis says, we can expect precedence to induce a kind of regularity: if the players manage to coordinate on one of the two equilibria, they should be expected to repeat the choices of actions that lead to that success, thus remaining in the same equilibrium.
Not only can we expect the players in such a repeated interaction to conform to a regularity, but also to expect others to do the same.
Their preference is to remain in a certain equilibrium given that others do too.
This summarizes Lewis's notion of a convention.

In order to extend this notion to \emph{linguistic} conventions, Lewis considers situations where the actions available involve sending and receiving signals or messages.
Thus, we could think of two players with different roles.
The first player, which Lewis calls the communicator~\parencite*[130]{lewis_convention_1969} and we will here call the sender, has knowledge about which of a number of possible states of affairs obtains, and depending on this information chooses a signal to send.
The second player, which Lewis calls the audience~\parencite*[130]{lewis_convention_1969} and we will here call the receiver, has knowledge about which signal the sender chose, and based on this information chooses one of several possible responses.
A preference relation exists between responses and states of affairs, and a payoff is attributed to each player based on the choices of both.
Note that Lewis assumes that no player has any preference regarding the particular signal that is used, provided that it enables coordination.
Formally, in order to describe the setup all we need is to specify a set of possible states of affairs $T$, a set of available signals or messages $M$, a set of responses or actions $A$, and the utility function $U : T \times A \rightarrow \mathbb{R}^2$.

Despite the added dimension of the signal exchange, these so-called signaling problems can be seen as particular cases of coordination problems if we consider the players' choices to be of contingency plans or strategies.
A communicator's contingency plan, or sender strategy as we will call it, is a specification of a choice of message for each possible state of affairs.
It thus describes the sender's behavior conditional on the state of affairs that obtains.
An audience's contingency plan, or receiver strategy as we will call it, analogously specifies a choice of action for each possible message.
Thus, formally, what the sender chooses is a function $\sigma : T \rightarrow M$ and the receiver a function $\rho : M \rightarrow A$.
The expected utility $EU$ of a pair of strategies $(\sigma,\rho)$ can be calculated using the utility function between states of affairs and actions as a sum of the payoffs for all cases, \emph{i.e.}:
$$
EU(\sigma, \rho) = \sum_{t \in T} U(t, \rho(\sigma(t)))
$$
As an example, consider a game with $T = \lbrace t_1, t_2 \rbrace$, $M = \lbrace m_1, m_2 \rbrace$, $A = \lbrace a_1, a_2 \rbrace$, and the following utility matrix:
\begin{center}
\begin{tabular}{r|c|c|}
\multicolumn{1}{r}{}
 & \multicolumn{1}{c}{$a_1$}
 & \multicolumn{1}{c}{$a_2$} \\ \cline{2-3}
   $t_1$ & $1,1$ & $0,0$ \\ \cline{2-3}
   $t_2$ & $0,0$ & $1,1$ \\ \cline{2-3}
\end{tabular}
\end{center}
Possible sender and receiver strategies could be, for example, $\sigma = \lbrace t_1 \mapsto m_2, t_2 \mapsto m_1 \rbrace$ and $\rho = \lbrace m_1 \mapsto a_2, m_2 \mapsto a_1 \rbrace$.
According to our formula above, these would have an expected utility of $2$ for both sender and receiver.
% the following expected utility:
% \begin{equation}
% \begin{split}
% EU(\sigma, \rho) & = \sum_{t \in T} U(t, \rho(\sigma(t))) \\
%                  & = U(t_1, \rho(\sigma(t_1))) + U(t_2, \rho(\sigma(t_2))) \\
%                  & = U(t_1, \rho(m_2)) + U(t_2, \rho(m_1)) \\
%                  & = U(t_1, t_1) + U(t_2, t_1) \\
%                  & = (1,1) + (1,1) \\
%                  & = (2,2)
% \end{split}
% \end{equation}
% Based on this, we could also create a matrix of expected utilities for this example as follows:
% \begin{center}
% \begin{tabular}{r|c|c|c|c|}
% \multicolumn{1}{r}{}
%  & \multicolumn{1}{c}{$m_1 \mapsto a_1$}
%  & \multicolumn{1}{c}{$m_1 \mapsto a_1$}
%  & \multicolumn{1}{c}{$m_1 \mapsto a_2$}
%  & \multicolumn{1}{c}{$m_1 \mapsto a_2$} \\
% \multicolumn{1}{r}{}
%  & \multicolumn{1}{c}{$m_2 \mapsto a_1$}
%  & \multicolumn{1}{c}{$m_2 \mapsto a_2$}
%  & \multicolumn{1}{c}{$m_2 \mapsto a_1$}
%  & \multicolumn{1}{c}{$m_2 \mapsto a_2$} \\ \cline{2-5}
%    $t_1 \mapsto m_1, t_2 \mapsto m_1$ & $1,1$ & $1,1$ & $1,1$ & $1,1$ \\ \cline{2-5}
%    $t_1 \mapsto m_1, t_2 \mapsto m_2$ & $1,1$ & $2,2$ & $0,0$ & $1,1$ \\ \cline{2-5}
%    $t_1 \mapsto m_2, t_2 \mapsto m_1$ & $1,1$ & $0,0$ & $2,2$ & $1,1$ \\ \cline{2-5}
%    $t_1 \mapsto m_2, t_2 \mapsto m_2$ & $1,1$ & $1,1$ & $1,1$ & $1,1$ \\ \cline{2-5}
% \end{tabular}
% \end{center}
They also represent one of the two stable conventions in this game, the other being the pair of strategies $\sigma = \lbrace t_1 \mapsto m_1, t_2 \mapsto m_2 \rbrace$ and $\rho = \lbrace m_1 \mapsto a_1, m_2 \mapsto a_2 \rbrace$.
Conventions of this kind in a signaling problem are what Lewis calls \emph{signaling systems}.

Brian Skyrms~\parencite*[80--104]{skyrms_evolution_1996} identifies some problems with the story so far.
Lewis' account of the stability of conventions rests on what could be considered strong demands for there to be the required common knowledge between the players involved.
Namely, there needs to be a state of affairs that indicates to everyone involved that a certain regularity will hold, as well as ``mutual ascription of some common inductive standards and background information, rationality, mutual ascription of rationality, and so on''~\parencite*[56--57]{lewis_convention_1969}.
These requirements can seem even more excessive if we consider how simple signaling systems are, compared to human languages.
The purpose of studying them was to tell a story about how language as a conventional system could get off the ground without any prior sort of agreement.
It is implausible to assume a high degree of rationality of the agents that started making use of primordial signaling systems that (hypothetically) evolved into languages.
Also, communication through simple message exchange is something that almost all animals do, from monkeys to birds to bees.
A plausible account of the origin of language should first explain how signaling systems like those could get started without a great deal of rationality from the part of the agents involved.

% One interesting aspect of these signaling systems is that, if we were to observe the repeated interactions of sender and receiver according to those strategies from a bird's eye view, we would be inclined to say that the signals are meaningful.
% For example, if we consider the latter example, a sender would always send $m_1$ when state of affairs $t_1$ obtains to which the receiver would always respond with action $a_1$ which maximizes both players' payoff.
% It is as if $m_1$ means either ``$t_1$ obtains'' or ``do $a_1$''.
%
